---
title: "Untitled"
author: "jp"
date: "6/5/2019"
output: html_document
params:
  origin: Sierra Leone Demo
  origin_login_file: sierra leone
  destination_login_file: aws_sierra
  cache: FALSE
---
# Setup 

load libraries...

```{r packages, echo = TRUE, results='hold', message=FALSE }

# list of required packages
    package.list = c("rlist", "knitr", "kableExtra", "xts", "leaflet", "RColorBrewer", "DT", "dygraphs", "httr", "jsonlite", "XML", "assertthat", "lubridate", "scales", "RcppRoll", "zoo", "gridExtra", "futile.logger", "tidyverse" )

# Function to test if package is installed 
    pkgTest <- function( package.list = package.list ){
        
        missing.packages = setdiff( package.list , rownames(installed.packages())) 
        if ( length( missing.packages ) > 0 ) install.packages( missing.packages ) 
    }


# Test if packages loaded
    pkgTest( package.list )

# load the packages
    lapply( package.list, suppressMessages( require ) , character.only = TRUE)


knitr::opts_chunk$set( echo = params$echo ,
                       fig.width =  8, 
                       # knitr.table.format = "html" ,
                       cache = params$cache # TRUE for testing; FALSE for production
                       )

```

and functions...

```{r sources }
source('../HMIS/DHIS2/dhis2-data-munging/bootstrap/bootstrap_functions.R')

source('../HMIS/DHIS2/dhis2_functions.R')

origin.folder = paste0( params$origin , "/"  )


 # login shortcuts
login_destination = function(){
    source( 
    paste0( origin.folder, 
            tolower( params$destination_login_file ) , "_login" )
    )

  loginDHIS2( baseurl, username, password)
}

login_destination()
```

Open Arc2 file and filter to SL

```{r}
      arc2_adm2_file =  "arc_adm2_rainfall.rds"  

       arc2_adm2 = read_rds( arc2_adm2_file )
       
       glimpse( arc2_adm2 )
       
       arc2_adm2.sl = arc2_adm2 %>% filter( NAME_0 %in% "Sierra Leone" ) %>%
         rename( districts = NAME_2 )


```

Get list of districts in ARC2 file

```{r}

   arc2.districts = arc2_adm2.sl %>% count( districts ) %>% select( -n )

    arc2.districts
    
```

Open list of SL districts

```{r}
 meta_data_file = paste0( params$origin , "/" ,
                           params$origin , "_metadata.rds" ) 

  if ( file.exists( meta_data_file) ){
   
      md = read_rds( meta_data_file )
  }

 district.level = md$organisationUnitLevels %>% filter( displayName %in% 'District' ) %>% pull(level )

 districts = md$organisationUnits %>%
   filter( level %in% district.level ) %>%
   select( id, name ) %>%
   rename( districts = name )
 
 districts

```

Link district names 

```{r}

 arc2.districts %>% left_join( districts , by = "districts")
```

Show unmatched

```{r}
 arc2.districts %>% anti_join( districts , by = "districts")
```

The arc dataset has "Western Area" instead of Western Rural and Western Urban.

Now, create a data element for ARC2 

```{r}
set.seed(94005004)
arc2.id = generateUID()

arc_de = tibble( name = "ARC2 mean rainfall" , shortName = "Mean Rain",
                 domainType = "AGGREGATE" , valueType = "NUMBER" , 
                 aggregationType = "SUM" ,
                 zeroIsSignificant = "true" ,
                 id = arc2.id )


 login_destination()

 url <- paste0( baseurl , 
                 "api/metadata?importStrategy=CREATE_AND_UPDATE&atomicMode=NONE"
                )
     

 de.list = list( dataElements = arc_de )

 post( url, de.list ) 
  
```


And, a new data element group, 'NASA'

<!-- TODO: FIX-- -->

```{r}
set.seed(94005005)
degroup.id = generateUID()



nasa_degp = tibble( name = "NASA" , 
                 id = degroup.id ,
                 dataElements = list(data.frame( id = arc2.id)) 
                 )



 login_destination()

 url <- paste0( baseurl , 
                 "api/metadata?importStrategy=CREATE_AND_UPDATE&atomicMode=NONE"
                )
     

   degrp.list = list( dataElementGroups = nasa_degp )
  
  post( url, degrp.list )
  
```

Now, prepare the dataset

```{r}
 data = arc2_adm2.sl %>% 
  left_join( districts , by = "districts") %>%
  filter( !is.na( id ) , !is.na( rain_mean ) ) %>%
  rename( orgUnit = id  ) %>%
  mutate( value = as.character( rain_mean ) ,
          dataElement = arc2.id ,
          # note: need to convert month to text and pad  so that always 2 characters
          period = paste0( year, str_pad( month , 2, pad = "0")  ) ) %>%
  select( orgUnit, dataElement ,  period , value )

glimpse( data )
```

And...

- After getting the data into shape, import it.

```{r import_data , echo = TRUE}

 
  url <- paste0( baseurl,
                 "api/dataValueSets?preheatCache=false&skipExistingCheck=false" )
  
# Import month by month
  periods = unique( data$period ) 
  periods = periods[ order( periods )]
  
for ( i in seq_along(periods) ){ 
  
  print( paste( periods[i] ) )
  
  d_out = data %>% 
    filter( period %in% periods[i] ) 
  
  p = POST( url, 
              body = toJSON( list( dataValues = d_out ) ,
                             auto_unbox = TRUE ) ,
              content_type_json()
    )
  
  # stop if did not import successfully
  status = fromJSON(content(p, "text" ))$description
  if ( !status %in% "Import process completed successfully" ){
    print( 'Import error')
    break
  } 
  
  # results
  print( fromJSON(content(p, "text" ))$importCount %>% data.frame() ) 
}

  
```

